stats-project

### 2020-12-02 R - Deliverable/MSDS 6371 - Project.Rmd
some changes

```{r message=FALSE, warning=FALSE}
library(olsrr)
library(tidyverse)
library(ggpubr)
library(ggplot2)
library(car)
library(data.table)
library(mltools)
```

```{r message=FALSE, warning=FALSE}
data = read_csv("../Data/train.csv")
data
#Convert <chr> to <fctr>
data = as.data.frame(unclass(data))
data
```

@@

# Select relevant columns and Neighborhood
train_data_q1 = data %>% select(Id, GrLivArea, Neighborhood, SalePrice ) %>% filter(Neighborhood == "NAmes" | Neighborhood ==  "Edwards" | Neighborhood ==  "BrkSide") %>% droplevels()
train_data_q1
train_data_q2 = data %>% select(Id, GrLivArea, Neighborhood, SalePrice ) 
train_data_q2
```

@@

press = function (model) {
  pr = resid(lm_fit_q1)/(1 - lm.influence(lm_fit_q1)$hat)
  return(sum(pr^2))
}

@@

```
#	ANALYSIS 1
```{r message=FALSE, warning=FALSE}
# One hot encode data
#one_encode_train_data_q1 =  one_hot(as.data.table(train_data_q1, dropUnusedLevels=TRUE))

#Complete Fit
# Neighborhood reference: Neighborhood_BrkSide
#lm_fit_q1 = lm(SalePrice ~. -Id-Neighborhood_BrkSide, data = one_encode_train_data_q1)

lm_fit_q1 = lm(SalePrice ~. -Id, data = train_data_q1)

press(lm_fit_q1)

@@

@@

RSS = c(crossprod(lm_fit_q1$residuals))
MSE = RSS / length(lm_fit_q1$residuals)
RMSE = sqrt(MSE)
RMSE

g1 = ggplot() + geom_histogram(aes(x=lm_fit_q1$residuals)) + ggtitle("Residuals Histogram")
g2 = ggplot(lm_fit_q1,aes(sample=lm_fit_q1$residuals)) + geom_qq() + geom_qq_line()  + ggtitle("Residuals QQ Plot")
g3 = ggplot(lm_fit_q1, aes(x=lm_fit_q1$fitted.values, y= lm_fit_q1$residuals)) + geom_point() +  geom_hline(yintercept=0) + xlab( "Predicted Value") + ylab("Residual") +  ggtitle("Residuals vs Predicted")
figure =  ggarrange(g1, g2, g3, ncol=3)
figure

# https://www.statmethods.net/stats/rdiagnostics.html

# Influence plot
influencePlot(lm_fit_q1, id.method="identify", main="Influence Plot", sub="Circle size is proportial to Cook's Distance" )

# Cook's D plot
cutoff = 4/((nrow(one_encode_train_data_q1)-length(lm_fit_q1$coefficients)-2))
plot(lm_fit_q1, which=4, cook.levels=cutoff)

#Durbin Watson Test for Colliniarity
durbinWatsonTest(lm_fit_q1)

# Outlier test
outlierTest(lm_fit_q1) # Bonferonni p-value for most extreme obs

# Multi-collinearity variance inflation factors
vif(lm_fit_q1)

@@

#	ANALYSIS 2
```{r message=FALSE, warning=FALSE}
# One hot encode data
one_encode_train_data_q2 =  one_hot(as.data.table(train_data_q2, dropUnusedLevels=TRUE))
one_encode_train_data_q2
# Neighborhood reference: Neighborhood_NAmes
lm_fit_q2 = lm(SalePrice ~. -Id -Neighborhood_NAmes, data = one_encode_train_data_q2)
summary(lm_fit_q2)
confint(lm_fit_q2,  level=0.995)
```
## Model Selection
### Stepwise
```{r message=FALSE, warning=FALSE}
#Stepwise Selection
stepwise_fit = ols_step_both_p(lm_fit_q2, pent =0.05, prem=0.05,  details = F)
stepwise_fit

```{r message=FALSE, warning=FALSE}
#Forward Selection
fw_fit = ols_step_forward_p(lm_fit_q2, penter=0.05, details = F)
fw_fit
```

```{r message=FALSE, warning=FALSE}
#Backward Selection
bw_fit = ols_step_backward_p(lm_fit_q2, penter=0.05, details = F)
bw_fit

RSS = c(crossprod(lm_fit_q2$residuals))
MSE = RSS / length(lm_fit_q2$residuals)
RMSE = sqrt(MSE)
RMSE

g1 = ggplot() + geom_histogram(aes(x=lm_fit_q2$residuals)) + ggtitle("Residuals Histogram")
g2 = ggplot(lm_fit_q2,aes(sample=lm_fit_q2$residuals)) + geom_qq() + geom_qq_line()  + ggtitle("Residuals QQ Pl0t")
g3 = ggplot(lm_fit_q2, aes(x=lm_fit_q2$fitted.values, y= lm_fit_q2$residuals)) + geom_point() +  geom_hline(yintercept=0) + xlab( "Predicted Value") + ylab("Residual") +  ggtitle("Residuals vs Predicted")
figure =  ggarrange(g1, g2, g3, ncol=3)
figure

# https://www.statmethods.net/stats/rdiagnostics.html

# Influence plot
influencePlot(lm_fit_q2, id.method="identify", main="Influence Plot", sub="Circle size is proportial to Cook's Distance" )

# Cook's D plot
cutoff = 4/((nrow(one_encode_train_data_q2)-length(lm_fit_q2$coefficients)-2))
plot(lm_fit_q2, which=4, cook.levels=cutoff)

#Durbin Watson Test for Colliniarity
durbinWatsonTest(lm_fit_q2)

# Outlier test
outlierTest(lm_fit_q2) # Bonferonni p-value for most extreme obs

# Multi-collinearity variance inflation factors
vif(lm_fit_q2)

>>>>

```{r message=FALSE, warning=FALSE}
library(olsrr)
library(tidyverse)
library(ggpubr)
library(ggplot2)
library(car)
library(data.table)
library(mltools)

# options
#options(scipen = 999)
```

```{r message=FALSE, warning=FALSE}
data_train= read_csv("../Data/train.csv")
data_test = read_csv("../Data/test.csv")
#Convert <chr> to <fctr>
data_train = as.data.frame(unclass(data_train))
data_test = as.data.frame(unclass(data_test))
data_test
```

@@

# Select relevant columns and Neighborhood
train_data_q1 = data_train %>% select(Id, GrLivArea, Neighborhood, SalePrice ) %>% filter(Neighborhood == "NAmes" | Neighborhood ==  "Edwards" | Neighborhood ==  "BrkSide") %>% droplevels()
test_data_q1 = data_test  %>% select(Id, GrLivArea, Neighborhood ) %>% filter(Neighborhood == "NAmes" | Neighborhood ==  "Edwards" | Neighborhood ==  "BrkSide") %>% droplevels()
train_data_q1
train_data_q2 = data_train %>% select(Id, GrLivArea, Neighborhood, SalePrice ) 
train_data_q2$GrLivArea2 = train_data_q2$GrLivArea^2

test_data_q2 = data_test %>% select(Id, GrLivArea, Neighborhood ) 
test_data_q2$GrLivArea2 = test_data_q2$GrLivArea^2

```{r}

lm_fit_experiment = lm(SalePrice ~. -Id-Utilities-Street-PoolQC-MiscFeature-Alley-Fence-FireplaceQu-LotFrontage-GarageCond-GarageFinish-GarageQual-GarageType-GarageYrBlt-BsmtExposure-BsmtFinType2-BsmtCond-BsmtFinType1-BsmtQual-MasVnrArea-MasVnrType-Electrical  , data = train)
lm_fit_experiment
#Neighborhood_Blmngtn

summary(lm_fit_experiment)
#confint(lm_fit_experiment,  level=0.995)

@@

press = function (model) {
  pr = resid(model)/(1 - lm.influence(model)$hat)
  return(sum(pr^2))
}

@@

```
#	ANALYSIS 1
```{r message=FALSE, warning=FALSE}
one_encode_train_data_q1 =  one_hot(as.data.table(train_data_q1, dropUnusedLevels=TRUE))

#Reference Neighborhood_BrkSide
lm_fit_q1 = lm(SalePrice ~. -Id-Neighborhood_BrkSide+Neighborhood_Edwards*GrLivArea+Neighborhood_NAmes*GrLivArea, data = one_encode_train_data_q1)

press(lm_fit_q1)
ols_press(lm_fit_q1)

@@

ols_plot_cooksd_bar(lm_fit_q1)
ols_plot_dffits(lm_fit_q1)
ols_plot_diagnostics(lm_fit_q1)

ols_plot_resid_fit(lm_fit_q1)
ols_plot_resid_pot(lm_fit_q1)
ols_pred_rsq(lm_fit_q1)
plot(lm_fit_q1)
```

```{r}
lm_fit_q1$residuals

@@

#	ANALYSIS 2
```{r message=FALSE, warning=FALSE}

one_encode_train_data_q2 =  one_hot(as.data.table(train_data_q2, dropUnusedLevels=TRUE))
#lm_fit_q2 = lm(SalePrice ~. -Id-Neighborhood_BrkSide, data = one_encode_train_data_q2)

#Neighborhood_Blmngtn

lm_fit_q2 = lm(SalePrice ~. -Id-Neighborhood_Blmngtn+(Neighborhood_BrkSide*GrLivArea+Neighborhood_Blueste+Neighborhood_BrDale+Neighborhood_ClearCr+Neighborhood_CollgCr+Neighborhood_Crawfor+Neighborhood_Edwards+Neighborhood_Gilbert+Neighborhood_IDOTRR+Neighborhood_MeadowV+Neighborhood_Mitchel+Neighborhood_NAmes+Neighborhood_NPkVill+Neighborhood_NridgHt+Neighborhood_NWAmes+Neighborhood_OldTown+Neighborhood_Sawyer+Neighborhood_SawyerW+Neighborhood_Somerst+Neighborhood_StoneBr+Neighborhood_SWISU+Neighborhood_Timber+Neighborhood_Veenker)*GrLivArea+(Neighborhood_BrkSide+Neighborhood_Blueste+Neighborhood_BrDale+Neighborhood_ClearCr+Neighborhood_CollgCr+Neighborhood_Crawfor+Neighborhood_Edwards+Neighborhood_Gilbert+Neighborhood_IDOTRR+Neighborhood_MeadowV+Neighborhood_Mitchel+Neighborhood_NAmes+Neighborhood_NPkVill+Neighborhood_NridgHt+Neighborhood_NWAmes+Neighborhood_OldTown+Neighborhood_Sawyer+Neighborhood_SawyerW+Neighborhood_Somerst+Neighborhood_StoneBr+Neighborhood_SWISU+Neighborhood_Timber+Neighborhood_Veenker)*GrLivArea2, data = one_encode_train_data_q2)

summary(lm_fit_q2)
confint(lm_fit_q2,  level=0.995)
```
## Model Selection
### Stepwise
```{r message=FALSE, warning=FALSE}
#Stepwise Selection
stepwise_fit = ols_step_both_p(lm_fit_q2, pent =0.05, prem=0.05,  details = F)
stepwise_fit
stepwise_model = stepwise_fit$model
ols_press(stepwise_model)
stepwise_model
summary(stepwise_model)
confint(stepwise_model)

ols_plot_cooksd_bar(stepwise_model)

stepwise_model_outliers = ols_plot_cooksd_bar(stepwise_model, print_plot=F)
stepwise_model_outliers

ols_plot_dffits(stepwise_model)
ols_plot_diagnostics(stepwise_model)

ols_plot_resid_fit(stepwise_model)
ols_plot_resid_pot(stepwise_model)
ols_pred_rsq(stepwise_model)
plot(stepwise_fit)
```

```{r message=FALSE, warning=FALSE}
#Forward Selection
fw_fit = ols_step_forward_p(lm_fit_q2, penter=0.05, details = F)
fw_fit
fw_model = fw_fit$model
ols_press(fw_model)
fw_model
summary(fw_model)
confint(fw_model)

ols_plot_cooksd_bar(fw_model)

fw_model_outliers = ols_plot_cooksd_bar(fw_model, print_plot=F)
fw_model_outliers

ols_plot_dffits(fw_model)
ols_plot_diagnostics(fw_model)

ols_plot_resid_fit(fw_model)
ols_plot_resid_pot(fw_model)
ols_pred_rsq(fw_model)
plot(fw_fit)
```

```{r message=FALSE, warning=FALSE}
#Backward Selection
bw_fit = ols_step_backward_p(lm_fit_q2, penter=0.05, details = F)
bw_fit
bw_model = bw_fit$model
ols_press(bw_model)
bw_model
summary(bw_model)
confint(bw_model)

ols_plot_cooksd_bar(bw_model)

bw_model_outliers = ols_plot_cooksd_bar(bw_model, print_plot=F)
bw_model_outliers

ols_plot_dffits(bw_model)
ols_plot_diagnostics(bw_model)

ols_plot_resid_fit(bw_model)
ols_plot_resid_pot(bw_model)
ols_pred_rsq(bw_model)
plot(bw_fit)

@@
